### 2016-03-21 manuscript.Rnw
Include reference to phangorn as core R phylo pkg

The R programming language is a popular tool for phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, and \texttt{RNeXML} \citep{Boettiger2015} each provide
functions to import and manipulate trees within R and save the results in
standard data formats. Additional packages including \texttt{phytools}
\citep{Revell2012}, \texttt{geiger} \citep{Pennell2014-geiger}, and
\texttt{ggtree} \citep{Yu-ggtree}, allow users to analyze and visualize data in
a phylogenetic context (see
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogenetic and comparative analyses, a growing number of R packages allows
users to query and access data from the web (e.g. \texttt{rFISHBASE}
\citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and \texttt{paleobioDB}
\citep{Varela2015}), such that data associated with taxa in a given phylogeny
can be obtained directly in R.

@@

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology. R greatly facilitates the analysis of large
datasets, and allows researchers to combine methods in novel ways because many
methods for comparative analyses are implemented, and because it is a relatively
easy to use programming language. Additionally, as more data are made available
online and accessible using APIs, several packages have been developed to
interact and download these datasets directly in R, thereby enabling direct and
reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general. For instance, the
rOpenSci-developed package \texttt{TreeBase} \citep{Boettiger2012TB} allows users
to access phylogenies stored in treeBASE (\url{www.treebase.org}).
\texttt{rotl} contributes to this initiative, and greatly extends the number of
taxa for which phylogenetic data can be retrieved within R, while allowing the
data from OTL to be combined with other sources easily.

>>>>

The R programming language is a popular tool for phylogenetics and comparative
analysis. The R packages \texttt{ape} \citep{Paradis2004}, \texttt{phylobase}
\citep{phylobase-0.8.0}, \texttt{phangorn} \citep{schliep2011} and \texttt{RNeXML} \citep{Boettiger2015} each provide
functions to import and manipulate trees within R and save the results in
standard data formats. Additional packages including \texttt{phytools}
\citep{Revell2012}, \texttt{geiger} \citep{Pennell2014-geiger}, and
\texttt{ggtree} \citep{Yu-ggtree}, allow users to analyze and visualize data in
a phylogenetic context (see
\url{https://cran.r-project.org/web/views/Phylogenetics.html} for a
comprehensive list of phylogenetics packages in R). In addition to packages for
phylogenetic and comparative analyses, a growing number of R packages allows
users to query and access data from the web (e.g. \texttt{rFISHBASE}
\citep{Boettiger2012}, \texttt{rAvis} \citep{Varela2014}, and \texttt{paleobioDB}
\citep{Varela2015}), such that data associated with taxa in a given phylogeny
can be obtained directly in R.

@@

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology and ecology. R greatly facilitates the 
analysis of large datasets, and allows researchers to combine methods in novel
ways because many methods for comparative analyses are implemented, and because
it is a relatively easy to use programming language. Additionally, as more data 
are made available online and accessible using APIs, several packages have been 
developed to interact and download these datasets directly in R, thereby 
enabling direct and reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general. For instance, the
rOpenSci-developed package \texttt{TreeBase} \citep{Boettiger2012TB} allows users
to access phylogenies stored in treeBASE (\url{www.treebase.org}).
\texttt{rotl} contributes to this initiative, and greatly extends the number of
taxa for which phylogenetic data can be retrieved within R, while allowing the
data from OTL to be combined with other sources easily.
